Channel Shaping Using Reconfigurable Intelligent Surfaces: From Diagonal to Beyond

### 2024-01-29 main.tex
:pencil: Add Pareto frontier problem

	Second, we propose a \gls{rcg} algorithm for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
	Specifically, blockwise update is performed along the geodesics\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} of the Stiefel manifold and evaluated compactly as matrix exponential.
	The proposed method features lower complexity and faster convergence than general manifold optimization \cite{Absil2009,Pan2022d}, and is used to solve the Pareto singular value problem.
	This is the first paper to tailor an efficient optimization framework that unleashes the design potential of asymmetric \gls{bd}-\gls{ris}.

@@

\begin{section}{\gls{bd}-\gls{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
	The \gls{bd}-\gls{ris} is modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that further divides into $G$ independent groups.
	Each group contains $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume the \gls{bd}-\gls{ris} features lossless asymmetric\footnote{While symmetric (impedance) network is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric passive components (e.g., ring hybrids and branch-line hybrids) may also be reconfigured in real time \cite{Ahn2006} and asymmetric \gls{bd}-\gls{ris} has been discussed in \cite{Li2023b,Li2023c,Bartoli2023}.} in-group connections without mutual coupling.
	The overall scattering matrix is thus block diagonal $\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G) \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$, where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix corresponding to group $g \in \mathcal{G} \triangleq \{1,\ldots,G\}$.
	Consider a quasi-static block fading model where $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$ denote the direct (transmitter-receiver), forward (transmitter-\gls{ris}), and backward (\gls{ris}-receiver) channels, respectively.
	The equivalent channel is
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels of \gls{ris} group $g$, respectively.
	\begin{remark}
		\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size 1 and $N_\mathrm{S}$, respectively.
	\end{remark}
	\begin{remark}
		Individual forward and backward \gls{csi} are generally required for \gls{bd}-\gls{ris} designs.
		This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
	\end{remark}
\end{section}

\begin{section}{Channel Singular Value Redistribution}
	\begin{subsection}{A Toy Example}
		% We first show the channel shaping advantage of \gls{bd}-\gls{ris} by a toy example.
		We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		For a fair comparison, we also consider symmetric impedance network with $\beta = \pi / 2$.
		% For a fair comparison, we also consider symmetric impedance network with $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		% also enforce symmetry with $\beta = \pi / 2$.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{Channel singular value shaping by diagonal and symmetry unitary \gls{ris}. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$. Direct link is absent.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		Interestingly, both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of angular parameters and scattering elements.
		A larger performance gap is expected for asymmetric \gls{bd}-\gls{ris}.
		This example emphasizes the importance of cooperative scattering and motivates further studies on channel shaping.

		We then analyze the channel shaping \emph{capability} of \gls{bd}-\gls{ris} under specific setups.
		\begin{subsubsection}{Rank-Deficient Channel}
			\label{sc:pc_rank_deficient}
			In rank-deficient channels, \gls{bd}-\gls{ris} $\mathbf{\Theta}^\mathrm{B}$ cannot achieve a higher \gls{dof} than diagonal \gls{ris} $\mathbf{\Theta}^\mathrm{D}$.
			This is because $\mathrm{sv}(\mathbf{\Theta}^\mathrm{B}) = \mathrm{sv}(\mathbf{\Theta}^\mathrm{D}) = \boldsymbol{1}$ and
			\begin{equation}
				\begin{split}
					\mathrm{rank}(\mathbf{H})
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \mathrm{rank}(\mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F}) \\
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \min \bigl( \mathrm{rank}(\mathbf{H}^\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}), \mathrm{rank}(\mathbf{H}^\mathrm{F}) \bigr).
				\end{split}
			\end{equation}
			% However, \gls{bd}-\gls{ris} can still guarantee a higher rate at finite \gls{snr}.
			Note \gls{bd}-\gls{ris} can still provide a higher indirect \gls{snr} as shown in Fig.~\ref{sm:pc_power_sx} and \ref{sm:pc_power_bond}.
		\end{subsubsection}

@@

		\begin{subsubsection}{Fully-Connected \glsentryshort{ris} Without Direct Link}
			Denote the singular value decomposition of forward / backward channels as $\mathbf{H}^{\mathrm{B}/\mathrm{F}} = \mathbf{U}^{\mathrm{B}/\mathrm{F}} \mathbf{\Sigma}^{\mathrm{B}/\mathrm{F}} {\mathbf{V}^{\mathrm{B}/\mathrm{F}}}^\mathsf{H}$.
			The composite channel is
			\begin{equation}
				\mathbf{H} = \mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F} = \mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H},
			\end{equation}
			where $\mathbf{X} = {\mathbf{V}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta} \mathbf{U}^\mathrm{F}$.
			% Since unitary matrices are closed under multiplication,
			\begin{proposition}
				In this case, the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with singular values $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$.
				% The only singular value bounds applied here are those apply to the singular value bound of $\mathbf{F}\mathbf{B}$
			\end{proposition}
			\begin{proof}
				We first observe that singular value control problem can be solved w.r.t. unitary $\mathbf{X}$ and retrieved by $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} \mathbf{X} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
				Also, $\mathrm{sv}(\mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\bar{\mathbf{U}}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{\bar{V}}{{}^\mathrm{B}}^\mathsf{H} \bar{\mathbf{U}}^\mathrm{F} \mathbf{\Sigma}^\mathrm{F} \mathbf{\bar{V}}{{}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\mathbf{BF})$ where $\bar{\mathbf{U}}^{\mathrm{B}/\mathrm{F}}$ and $\bar{\mathbf{V}}^{\mathrm{B}/\mathrm{F}}$ are arbitrary unitary matrices.
			\end{proof}
			The problem now becomes, given $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$, what can we say about the singular value of $\mathbf{BF}$.
			One comprehensive answer is Horn's inequality \cite{Bhatia2001}: for all admissible triples $(I, J, K)$,
			\begin{equation}
				\prod_{k \in {K}} \sigma_k(\mathbf{BF}) \leq \prod_{i \in {I}} \sigma_i(\mathbf{B}) \prod_{j \in {J}} \sigma_j(\mathbf{F}).
			\end{equation}
			It gives upper bound on the largest singular value and lower bound on the smallest singular value:
			\begin{align}
				\sigma_1(\mathbf{BF}) \le \sigma_1(\mathbf{B}) \sigma_1(\mathbf{F}) \label{iq:singular_value_largest}\\
				\sigma_N(\mathbf{BF}) \geq \sigma_N(\mathbf{B}) \sigma_N(\mathbf{F}) \label{iq:singular_value_smallest}.
			\end{align}
			Another useful result is introduced in \cite{Hogben2013}: for all $p>0$,
			\begin{equation}
				\sum_n \sigma_n^p(\mathbf{BF}) \leq \sum_n \sigma_n^p(\mathbf{B}) \sigma_n^p(\mathbf{F}) \label{iq:singular_value_power_sum}.
			\end{equation}
			When $p=2$, it implies the channel energy is upper bounded by the sum of element-wise power product of the forward and backward channels, as illustrated in Fig.~\subref*{sm:pc_power_bond_nd}.
			Interestingly, \eqref{iq:singular_value_largest}--\eqref{iq:singular_value_power_sum} are simultaneously tight when $\mathbf{X} = \mathbf{I}$ and $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
			This solution was claimed in \cite{Bartoli2023} to achieve channel capacity, but it is not true at moderate \gls{snr}.
		\end{subsubsection}

>>>>

	Second, we propose a \gls{rcg} algorithm for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
	Specifically, block-wise update is performed along the geodesics\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} and evaluated compactly by the exponential map.
	The proposed method features lower complexity and faster convergence than general manifold optimization \cite{Absil2009,Pan2022d}, and is used to solve the Pareto singular value problem.
	This is the first paper to tailor an efficient optimization framework that unleashes the design potential of asymmetric \gls{bd}-\gls{ris}.

@@

\begin{section}{\gls{bd}-\gls{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
	The \gls{bd}-\gls{ris} is modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that further divides into $G$ independent groups.
	Each group contains $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume the \gls{bd}-\gls{ris} features lossless asymmetric\footnote{While symmetric impedance network is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric passive components (e.g., ring hybrids and branch-line hybrids) may also be reconfigured in real time \cite{Ahn2006}. Asymmetric \gls{bd}-\gls{ris} has been discussed in \cite{Li2023b,Li2023c,Bartoli2023}.} in-group connections without mutual coupling.
	The overall scattering matrix is thus block diagonal $\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G) \in \mathbb{U}^{N_\mathrm{S} \times N_\mathrm{S}}$, where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix corresponding to group $g \in \mathcal{G} \triangleq \{1,\ldots,G\}$.
	Consider a quasi-static block fading model where $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$ denote the direct (transmitter-receiver), forward (transmitter-\gls{ris}), and backward (\gls{ris}-receiver) channels, respectively.
	The equivalent channel is
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels of \gls{ris} group $g$, respectively.
	\begin{remark}
		\gls{bd}-\gls{ris} reduces to diagonal \gls{ris} and unitary \gls{ris} with group size 1 and $N_\mathrm{S}$, respectively.
	\end{remark}
	\begin{remark}
		Individual forward and backward \gls{csi} are required for \gls{bd}-\gls{ris} designs.
		This is different from diagonal \gls{ris} where estimating their product is usually sufficient.
		Later we will show the potential benefits from the \gls{csi} overhead.
	\end{remark}
\end{section}

\begin{section}{Channel Singular Value Redistribution}
	\begin{subsection}{A Toy Example}
		We first illustrate the channel shaping capabilities of different \gls{ris} by a toy example.
		% We first use a toy example to illustrate that \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		Consider a $2 \times 2 \times 2$ setup where the direct link is blocked.
		The diagonal \gls{ris} is modeled by $\mathbf{\Theta}_\mathrm{D} = \mathrm{diag}(e^{\jmath \theta_1}, e^{\jmath \theta_2})$ while the unitary \gls{bd}-\gls{ris} has 4 independent angular parameters
		\begin{equation}
			\mathbf{\Theta}_\mathrm{U} = e^{\jmath \phi} \begin{bmatrix}
				e^{\jmath \alpha} \cos \psi  & e^{\jmath \beta} \sin \psi   \\
				-e^{-\jmath \beta} \sin \psi & e^{-\jmath \alpha} \cos \psi
			\end{bmatrix}.
		\end{equation}
		In particular, $\phi$ has no impact on the singular value because $\mathrm{sv}(e^{\jmath \phi} \mathbf{A}) = \mathrm{sv}(\mathbf{A})$.
		For a fair comparison, we also consider symmetric impedance network with $\beta = \pi / 2$.
		% For a fair comparison, we also consider symmetric impedance network with $\beta = \pi / 2$ such that both architectures have the same number of angular parameters.
		% also enforce symmetry with $\beta = \pi / 2$.
		\begin{figure}
			\centering
			\includegraphics[width=\columnwidth]{assets/simulation/pc_singular_toy.eps}
			\caption{Channel singular value shaping by diagonal and symmetry unitary \gls{ris}. $(N^\mathrm{T}, N^\mathrm{S}, N^\mathrm{R}) = (2, 2, 2)$. Direct link is absent.}
			\label{sm:pc_singular_toy}
		\end{figure}
		Fig.~\ref{sm:pc_singular_toy} shows the channel singular values achieved by an exhaustive grid search over $(\theta_1, \theta_2)$ for diagonal \gls{ris} and $(\alpha, \psi)$ for symmetric unitary \gls{ris}.
		Interestingly, both singular values can be manipulated up to $9\%$ using diagonal \gls{ris} and $42\%$ using symmetric \gls{bd}-\gls{ris}, despite both architectures have the same number of angular parameters and scattering elements.
		A larger performance gap is expected when asymmetric \gls{bd}-\gls{ris} is available.
		This example shows \gls{bd}-\gls{ris} can provide a wider dynamic range of channel singular values.
		It emphasizes the importance of cooperative scattering and motivates further studies on channel shaping.
	\end{subsection}

	\begin{subsection}{Pareto Frontier Characterization}
		We then characterize the Pareto frontier of channel singular values by maximizing their weighted sum
		\begin{maxi!}
			{\scriptstyle{\mathbf{\Theta}}}{\sum_n \rho_n \sigma_n(\mathbf{H})}{\label{op:pc_singular}}{\label{ob:pc_singular}}
			\addConstraint{\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g=\mathbf{I}, \quad \forall g,}{}{}
		\end{maxi!}
		where $n \in \{1,\ldots,\min(N_\mathrm{T}, N_\mathrm{R})\}$ and $\rho_n$ is the weight of the $n$-th singular value that can be positive, zero, or negative.
		Varying $\{\rho_n\}$ allows us to explore the entire achievable singular value region.
		Problem~\eqref{op:pc_singular} thus generalizes all relevant metrics and provides a powerful channel shaping framework.
		For group $g$, the objective is smooth in $\mathbf{\Theta}_g$ and the feasible domain corresponds to the Stiefel manifold.
		% Those properties motivate us to propose a \gls{rcg} algorithm for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
		% Those motivate us to propose a general \gls{rcg} algorithm for asymmetric \gls{bd}-\gls{ris} with arbitrary group size, which will be detailed in Section~\ref{sc:rcg}.
		Those properties motivate us to zoom out to general smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.

		Inspired by \cite{Abrudan2008,Abrudan2009}, we propose a block-wise \gls{rcg} update along the geodesics on the Lie group of unitary matrices $\mathbb{U}^{L \times L}$.
		The gradient

		% It effectively solves general smooth asymmetric \gls{bd}-\gls{ris} problems with arbitrary group size.
		% The steps are summarized in Algorithm~\ref{ag:rcg}, where line \ref{ln:gradient_euclidean} computes the gradient in the Euclidean space, line \ref{ln:gradient_riemannian} translates it to the Riemannian manifold,
		% \footnote{The geodesics of the Stiefel manifold have compact expressions described by exponential map.}

% \footnote{The geodesics of the Stiefel manifold have compact expressions described by exponential map.}
		\setalgorithmcaptionfont{\small}
		\begin{algorithm}[!t]
			\small
			\caption{Block-wise \gls{rcg} update for asymmetric \gls{bd}-\gls{ris}}
			\label{ag:rcg}
			\begin{algorithmic}[1]
				\Require $f(\mathbf{\Theta})$, $G$
				\Ensure $\mathbf{\Theta}^\star$
				\State $r \gets 0$, $\mathbf{\Theta}^{(0)}$
				\Repeat
					\State $r \gets r+1$
					\For {$g \gets 1$ to $G$}
						\State $\nabla_{\mathrm{E},g}^{(r)} \gets {\partial f(\mathbf{\Theta}_g^{(r)})}/{\partial \mathbf{\Theta}_g^*}$\label{ln:gradient_euclidean}
						\State $\nabla_{\mathrm{R},g}^{(r)} \gets \nabla_{\mathrm{E},g}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {\nabla_{\mathrm{E},g}^{(r)}}^\mathsf{H}$\label{ln:gradient_riemannian}
						\State $\gamma_g^{(r)} \gets {\mathrm{tr}\bigl((\nabla_{\mathrm{R},g}^{(r)} - \nabla_{\mathrm{R},g}^{(r-1)}) {\nabla_{\mathrm{R},g}^{(r)}}^\mathsf{H}\bigr)}/{\mathrm{tr}\bigl(\nabla_{\mathrm{R},g}^{(r-1)} {\nabla_{\mathrm{R},g}^{(r-1)}}^\mathsf{H}\bigr)}$\label{ln:parameter_polak_approx}
						\State $\mathbf{D}_g^{(r)} \gets \nabla_{\mathrm{R},g}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)}$\label{ln:direction_conjugate}
						\State $\mu_g^{(r)} \gets \arg\max_{\mu_g} f\bigl(\exp\bigl(\mu_g \mathbf{D}_g^{(r)}\bigr) \mathbf{\Theta}_g^{(r)}\bigr)$\label{ln:step_armijo}
						\State $\mathbf{\Theta}_g^{(r+1)} \gets \exp\bigl(\mu_g^{(r)} \mathbf{D}_g^{(r)}\bigr) \mathbf{\Theta}_g^{(r)}$\label{ln:update_geodesic}
					\EndFor
				\Until $\lvert R^{(r)} - R^{(r-1)} \rvert / R^{(r-1)} \le \epsilon$
			\end{algorithmic}
		\end{algorithm}

		% fun = @(Theta_g) rho' * svd(H_d + H_b(:, S) * Theta_g * H_f(S, :) + H_b(:, S_c) * Theta(S_c, S_c) * H_f(S_c, :));
		% G_e(S, S) = gradient_euclidean(H, H_f(S, :), H_b(:, S), rho);
		% G_r(S, S) = gradient_riemannian(Theta(S, S), G_e(S, S));
		% D(S, S) = direction_conjugate(G_r(S, S), struct('G_r', iter.G_r(S, S), 'D', iter.D(S, S), 'counter', iter.counter));
		% Theta(S, S) = step_armijo(fun, Theta(S, S), D(S, S));

		% we update $\mathbf{\Theta}$ by \gls{rcg} method along the geodesics.
		% \begin{remark}
		% 	A geodesic refers to the shortest path between two points in a Riemannian manifold.
		% 	% For unitary constraint \eqref{cs:pc_rate_unitary}, the
		% 	Unitary constraint \eqref{cs:pc_rate_unitary} translates to a Stiefel manifold where the geodesics have simple expressions described by the exponential map \cite{Abrudan2008}.
		% 	% Geodesics are the locally shortest paths in a Riemannian manifold.
		% 	% A Geodesics are the locally shortest paths in a Riemannian manifold.
		% 	% on Riemannian manifold are the shortest path between two points.
		% \end{remark}
		For general optimization problems with block unitary constraint, the adapted \gls{rcg} method at iteration $r$ for block $g$ is summarized below, where $f\bigl(\mathbf{\Theta}_g^{(r)}\bigr)$ is the objective function also evaluated over $\bigl\{ \{\mathbf{\Theta}_{g'}^{(r+1)}\}_{g'<g}, \{\mathbf{\Theta}_{g'}^{(r)}\}_{g'>g} \bigr\}$.
		\begin{enumerate}
			\item Compute the Euclidean gradient
			\begin{equation}
				{\nabla_g^\mathrm{E}}^{(r)} = \frac{\partial f\bigl(\mathbf{\Theta}_g^{(r)}\bigr)}{\partial \mathbf{\Theta}_g^*};
				\label{eq:rcg_gradient_euclidean}
			\end{equation}
			\item Translate to the Riemannian gradient
			\begin{equation}
				{\nabla_g^\mathrm{R}}^{(r)} = {\nabla_g^\mathrm{E}}^{(r)} {\mathbf{\Theta}_g^{(r)}}^\mathsf{H} - \mathbf{\Theta}_g^{(r)} {{\nabla_g^\mathrm{E}}^{(r)}}^\mathsf{H};
				\label{eq:rcg_gradient_riemannian}
			\end{equation}
			\item Determine the weight factor
			\begin{equation}
				\gamma_g^{(r)} = \frac{\mathrm{tr}\Bigl(\bigl({\nabla_g^\mathrm{R}}^{(r)} - {\nabla_g^\mathrm{R}}^{(r-1)}\bigr) {{\nabla_g^\mathrm{R}}^{(r)}}^\mathsf{H}\Bigr)}{\mathrm{tr}\Bigl({\nabla_g^\mathrm{R}}^{(r-1)} {{\nabla_g^\mathrm{R}}^{(r-1)}}^\mathsf{H}\Bigr)};
				\label{eq:rcg_weight}
			\end{equation}
			\item Compute the conjugate direction
			\begin{equation}
				\mathbf{D}_g^{(r)} = {\nabla_g^\mathrm{R}}^{(r)} + \gamma_g^{(r)} \mathbf{D}_g^{(r-1)};
				\label{eq:rcg_direction}
			\end{equation}
			\item Determine the Armijo step size\footnote{To double the step size, simply square the argument instead of recomputing the matrix exponential, i.e., $\exp(2 \mu_g \mathbf{D}_g) = \exp^2(\mu_g \mathbf{D}_g)$.}
			\begin{equation}
				\mu_g^{(r)} = \arg\max_{\mu_g} f\Bigl(\exp\bigl(\mu_g \mathbf{D}_g^{(r)}\bigr) \mathbf{\Theta}_g^{(r)}\Bigr);
				\label{eq:rcg_step}
			\end{equation}
			\item Perform rotational update along local geodesics
			\begin{equation}
				\mathbf{\Theta}_g^{(r+1)} = \exp\Bigl(\mu_g^{(r)} \mathbf{D}_g^{(r)}\Bigr) \mathbf{\Theta}_g^{(r)}.
				\label{eq:rcg_update}
			\end{equation}
		\end{enumerate}

	\end{subsection}

	\begin{subsection}{Analytical Bounds}
		We then analyze the channel shaping \emph{capability} of \gls{bd}-\gls{ris} under specific setups.
		\begin{subsubsection}{Rank-Deficient Channel}
			\label{sc:pc_rank_deficient}
			In rank-deficient channels, \gls{bd}-\gls{ris} $\mathbf{\Theta}^\mathrm{B}$ cannot achieve a higher \gls{dof} than diagonal \gls{ris} $\mathbf{\Theta}^\mathrm{D}$.
			This is because $\mathrm{sv}(\mathbf{\Theta}^\mathrm{B}) = \mathrm{sv}(\mathbf{\Theta}^\mathrm{D}) = \boldsymbol{1}$ and
			\begin{equation}
				\begin{split}
					\mathrm{rank}(\mathbf{H})
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \mathrm{rank}(\mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F}) \\
					& \le \mathrm{rank}(\mathbf{H}^\mathrm{D}) + \min \bigl( \mathrm{rank}(\mathbf{H}^\mathrm{B}), \mathrm{rank}(\mathbf{\Theta}), \mathrm{rank}(\mathbf{H}^\mathrm{F}) \bigr).
				\end{split}
			\end{equation}
			% However, \gls{bd}-\gls{ris} can still guarantee a higher rate at finite \gls{snr}.
			Note \gls{bd}-\gls{ris} can still provide a higher indirect \gls{snr} as shown in Fig.~\ref{sm:pc_power_sx} and \ref{sm:pc_power_bond}.
		\end{subsubsection}

@@

		\begin{subsubsection}{Fully-Connected \glsentryshort{ris} Without Direct Link}
			Denote the singular value decomposition of forward / backward channels as $\mathbf{H}^{\mathrm{B}/\mathrm{F}} = \mathbf{U}^{\mathrm{B}/\mathrm{F}} \mathbf{\Sigma}^{\mathrm{B}/\mathrm{F}} {\mathbf{V}^{\mathrm{B}/\mathrm{F}}}^\mathsf{H}$.
			The composite channel is
			\begin{equation}
				\mathbf{H} = \mathbf{H}^\mathrm{B} \mathbf{\Theta} \mathbf{H}^\mathrm{F} = \mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H},
			\end{equation}
			where $\mathbf{X} = {\mathbf{V}^\mathrm{B}}^\mathsf{H} \mathbf{\Theta} \mathbf{U}^\mathrm{F}$.
			% Since unitary matrices are closed under multiplication,
			\begin{proposition}
				In this case, the singular value bounds on $\mathbf{H}$ are equivalent to the singular value bounds on $\mathbf{BF}$, where $\mathbf{B}$ and $\mathbf{F}$ are arbitrary matrices with singular values $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$.
				% The only singular value bounds applied here are those apply to the singular value bound of $\mathbf{F}\mathbf{B}$
			\end{proposition}
			\begin{proof}
				We first observe that singular value control problem can be solved w.r.t. unitary $\mathbf{X}$ and retrieved by $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} \mathbf{X} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
				Also, $\mathrm{sv}(\mathbf{U}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{X} \mathbf{\Sigma}^\mathrm{F} {\mathbf{V}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\bar{\mathbf{U}}^\mathrm{B} \mathbf{\Sigma}^\mathrm{B} \mathbf{\bar{V}}{{}^\mathrm{B}}^\mathsf{H} \bar{\mathbf{U}}^\mathrm{F} \mathbf{\Sigma}^\mathrm{F} \mathbf{\bar{V}}{{}^\mathrm{F}}^\mathsf{H}) = \mathrm{sv}(\mathbf{BF})$ where $\bar{\mathbf{U}}^{\mathrm{B}/\mathrm{F}}$ and $\bar{\mathbf{V}}^{\mathrm{B}/\mathrm{F}}$ are arbitrary unitary matrices.
			\end{proof}
			The problem now becomes, given $\mathbf{\Sigma}^\mathrm{B}$ and $\mathbf{\Sigma}^\mathrm{F}$, what can we say about the singular value of $\mathbf{BF}$.
			One comprehensive answer is Horn's inequality \cite{Bhatia2001}: for all admissible triples $(I, J, K)$,
			\begin{equation}
				\prod_{k \in {K}} \sigma_k(\mathbf{BF}) \leq \prod_{i \in {I}} \sigma_i(\mathbf{B}) \prod_{j \in {J}} \sigma_j(\mathbf{F}).
			\end{equation}
			It gives upper bound on the largest singular value and lower bound on the smallest singular value:
			\begin{align}
				\sigma_1(\mathbf{BF}) \le \sigma_1(\mathbf{B}) \sigma_1(\mathbf{F})\label{iq:singular_value_largest}\\
				\sigma_N(\mathbf{BF}) \geq \sigma_N(\mathbf{B}) \sigma_N(\mathbf{F})\label{iq:singular_value_smallest}.
			\end{align}
			Another useful result is introduced in \cite{Hogben2013}: for all $p>0$,
			\begin{equation}
				\sum_n \sigma_n^p(\mathbf{BF}) \leq \sum_n \sigma_n^p(\mathbf{B}) \sigma_n^p(\mathbf{F})\label{iq:singular_value_power_sum}.
			\end{equation}
			When $p=2$, it implies the channel energy is upper bounded by the sum of element-wise power product of the forward and backward channels, as illustrated in Fig.~\subref*{sm:pc_power_bond_nd}.
			Interestingly, \eqref{iq:singular_value_largest}--\eqref{iq:singular_value_power_sum} are simultaneously tight when $\mathbf{X} = \mathbf{I}$ and $\mathbf{\Theta} = \mathbf{V}^\mathrm{B} {\mathbf{U}^\mathrm{F}}^\mathsf{H}$.
			This solution was claimed in \cite{Bartoli2023} to achieve channel capacity, but it is not true at moderate \gls{snr}.
		\end{subsubsection}
